### 2015-12-23 manuscript.tex
initial commit

\documentclass[10pt,a4paper]{article}

%% packages
\usepackage{color}
\usepackage{pgf}
\usepackage[margin=1in]{geometry}
\usepackage[normalem]{ulem}
\usepackage{cite}
\usepackage{natbib}
\usepackage{authblk}
\usepackage{listings}
\usepackage{color}
\usepackage[colorlinks=true,
            linkcolor=red,
            urlcolor=blue,
            citecolor=gray]{hyperref}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=bash,
  frame=none,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\newcommand{\X}{\textcolor{red}{\bf X}}

\newcommand{\og}{\emph{O. glumaepatula }}

\title{More than one Author with different Affiliations}
\author[1,$\dag$]{Arun Durvasula}
\author[2,$\dag$]{Paul J. Hoffman}
\author[1]{Tyler V. Kent}
\author[2]{Chaochih Liu}
\author[2]{Thomas J. Y. Kono}
\author[2]{Peter L. Morrell}
\author[1,3,*]{Jeffrey Ross-Ibarra}
\affil[1]{Department of Plant Sciences, University of California, Davis, CA 95616}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN 55108}
\affil[3]{Center for Population Biology and Genome Center, University of California, Davis, CA 95616}
\affil[$\dag$]{These authors contributed equally.}
\affil[*]{email: rossibarra@ucdavis.edu}

\renewcommand\Authands{ and }

\begin{document}

\title{ANGSD-wrapper: utilities for analyzing next generation sequencing data}
\maketitle

\begin{abstract}
High throughput sequencing has changed many aspects of population genetics, molecular ecology, and related fields, affecting both experimental design and data analysis.
The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data. 
The package is specifically designed to produce more accurate results for samples with low sequencing depth, but it handles a wide array of sampling and experimental designs and makes use of full genome data.
Here we present ANGSD-wrapper, a user-friendly interface to ANGSD.
ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses. ANGSD-wrapper also provides interactive graphing of ANGSD results, thus enhancing data exploration.
We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
\end{abstract}

\section*{Introduction}

High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or survey fine-scale patterns of variation along the genome.  
Application of these methods has led to changes in both experimental design and data analysis \citep{ekblom2011applications}.
Many of the popular software packages used by researchers \citep[see][]{excoffier2006computer}  were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated. A particular challenge with short read sequencing has been higher per base pair rates of errors and missing data.

A number of tools have recently been published to handle high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}, but the majority of these either make limiting assumptions about the data (e.g., all sites have been sequenced, all genomes are haploid, sequencing is to sufficient depth, all individuals are outcrossing) or are specialized tools offering a narrow set of analysis options. 
Korneliussen et al. \citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including calculating the ABBA BABA D-statistic, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}
One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.

Here we present ANGSD-wrapper, a user-friendly interface to ANGSD. 
ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines inherent in ANGSD as well as pipelines involving related programs such as ngsPopGen, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
Because the large volume of data associated with high throughput sequence analysis is often difficult to explore by hand, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.  
We demonstrate some of the analyses possible using ANGSD-wrapper in an initial exploration of low-coverage whole-genome sequence data from populations of the endangered wild rice species \og  and compare to the domesticated \emph{Oryza sativa}.
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.

\section*{Methods}
ANGSD-wrapper is a set of configuration files and scripts written in the Bash UNIX shell.  
The scripts can be run either on a standalone computer with a UNIX terminal, or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.  
An installation of the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.  
The visualization portion of ANGSD-wrapper also requires installation of the R packages shiny \citep{shiny}, genomeIntervals \citep{genomeIntervals}, and ape \citep{APE}.

ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. ANGSD-wrapper provides a common configuration file, common.conf, which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
Each analysis is run using a simple command:

\begin{lstlisting}
$ angsd-wrapper <method> <configuration_file>
\end{lstlisting}

A detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.

The visualization software included with ANGSD-wrapper is contained within
it's own directory called shinyGraphing.
This application must be started in R and can be accessed locally from a web browser. This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
Each tab in the GUI contains plots for different ANGSD methods.

In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate file of results. 
The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.

%\begin{figure}
%\centering
%\includegraphics[width=130mm]{theta_rice_chr12.png}
%\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across chromosome 12 of \og using ANGSD-wrapper\label{fig:theta} }%\pjh{capitalized 'ANGSD' in ANGSD-wrapper}
%\end{figure}

\section*{Results}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
\caption{Example analysis workflow diagram.}
\label{fig:workflow}
\end{figure}

\section*{Discussion}

\section*{Conclusions}
Our software ANGSD-wrapper provides an intuitive and easy-to-use interface to employ the powerful and flexible suite of population genetic analyses developed in ANGSD \citep{korneliussen2014angsd} and permits the exploration of genome-scale results through interactive visualization.
ANGSD-wrapper is under active development, and work is already underway on a new release that will incorporate updates to the ANGSD software package.

\section*{Acknowledgements}
We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM) and funding from the UC Davis Plant Sciences Department.  We thank members of the Ross-Ibarra and Morrell labs for discussion and software testing. We thank Eric Fuchs for access to the test dataset. We thank the authors of ANGSD and related programs for answering questions, particularly Matteo Fumagalli and Filipe Vieira. Finally, we would like to thank Felix Andrews for statistical advice, although we did not follow it.

\bibliographystyle{unsrt}
\bibliography{references}

\input{supp.tex}

\end{document}

### 2015-12-28 cover_letter.tex
adding to cover letter

ANGSD-wrapper is intended to simplify analysis and visualization of next generation resequencing data.

Visualization of analysis results is implemented using Shiny graphics in R.

The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.

### 2016-01-13 manuscript.tex
added new screenshot
added fig3 and admixture
figure label changes
edits

@@
 
+\usepackage[colorinlistoftodos]{todonotes}
+\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+%\setlength{\marginparwidth}{1.5cm}
+\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
+
 \newcommand{\X}{\textcolor{red}{\bf X}}
-
-\newcommand{\og}{\emph{O. glumaepatula }}
 
@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data. 
-The package is specifically designed to produce more accurate results for samples with low sequencing depth, but it handles a wide array of sampling and experimental designs and makes use of full genome data.
-Here we present ANGSD-wrapper, a user-friendly interface to ANGSD.
-ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
-ANGSD-wrapper also provides interactive graphing of ANGSD results, thus enhancing data exploration.
+The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
+Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
+ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses and provides interactive graphing of ANGSD results to enhance data exploration.
 We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
@@
 Many of the popular software packages used by researchers \citep[see][]{excoffier2006computer}  were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated.
-A particular challenge with short read sequencing has been higher per base pair rates of errors and missing data.
+In particular, short read sequencing has brought new challenges, including highly variable coverage, missing data, and high per-nucleotide error rates.
 
@@
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including calculating the ABBA BABA D-statistic, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}
+\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, Patterson's D-statistic test of admixture \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.
 One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
@@
 Because the large volume of data associated with high throughput sequence analysis is often difficult to explore by hand, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
-We demonstrate some of the analyses possible using ANGSD-wrapper in an initial exploration of low-coverage whole-genome sequence data from populations of the endangered wild rice species \og  and compare to the domesticated \emph{Oryza sativa}.
+We demonstrate some of the analyses possible using ANGSD-wrapper using low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
 ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
@@
 
-A detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
+Analyses supported by ANGSD-wrapper are shown in table \X, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.\jri{add a simple table of all analyses supported.}
 
-The visualization software included with ANGSD-wrapper is contained within
-it's own directory called shinyGraphing.
+The visualization software included with ANGSD-wrapper is contained within it's own directory called shinyGraphing.
 This application must be started in R and can be accessed locally from a web browser.
@@
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
+\jri{is this really chromosome 10 of maize?
+looks like only first 10Mb.
+if so, should mention that in figure legend}
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=130mm]{theta_rice_chr12.png}
-%\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across chromosome 12 of \og using ANGSD-wrapper\label{fig:theta} }%\pjh{capitalized 'ANGSD' in ANGSD-wrapper}
-%\end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=130mm]{figures/ThetasPlot.png}
+\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across chromosome 10 of {\it Zea mays} spp. {\it mays} using ANGSD-wrapper\label{fig:theta} }
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/figure3big.pdf}
+\caption{Summary statistics for {\it Zea mays}. Site frequency spectra for A. maize and B. teosinte. C. distribution of pairwise differences for maize (blue) and teosinte (orange).}
+\label{fig:figure3}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/mt4labeled.png}
+\caption{Admixture analysis for {\it Zea mays} spp. {\it mays}, {\it Zea mays} spp. {\it parviglumas}, and {\it Zea mays} spp. {mexicana} (Mex.) with K=4 source populations. }
+\label{fig:admixture}
+\end{figure}
+
 \section*{Discussion}
@@
 Our software ANGSD-wrapper provides an intuitive and easy-to-use interface to employ the powerful and flexible suite of population genetic analyses developed in ANGSD \citep{korneliussen2014angsd} and permits the exploration of genome-scale results through interactive visualization.
-ANGSD-wrapper is under active development, and work is already underway on a new release that will incorporate updates to the ANGSD software package.
+ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.
 
@@
 We thank members of the Ross-Ibarra and Morrell labs for discussion and software testing.
-We thank Eric Fuchs for access to the test dataset.
 We thank the authors of ANGSD and related programs for answering questions, particularly Matteo Fumagalli and Filipe Vieira.

### 2016-01-13 manuscript.tex
added table and addressed fig 1 comment

@@
 
+\begin{table}
+\begin{center}
+    \begin{tabular}{ | l | p{5cm} |}
+    \hline
+    Methods Implemented & Interactive Graphing \\ \hline
+    ABBA BABA & Yes \\ \hline
+    Admixture & Yes  \\ \hline
+    Ancestral Sequence & N/A \\ \hline
+    Genotype Likelihoods & N/A \\ \hline
+    Inbreeding Coefficients & N/A \\ \hline
+    PCA & Yes \\ \hline
+    Site Frequency Spectrum & Yes \\ \hline
+    Theta Calculation & Yes \\ \hline
+    \end{tabular}
+    \caption{Table of methods implemented in ANGSD-wrapper}
+    \label{tab:methods}
+    \end{center}
+\end{table}
+
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
@@
 
-Analyses supported by ANGSD-wrapper are shown in table \X, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.\jri{add a simple table of all analyses supported.}
+Analyses supported by ANGSD-wrapper are shown in table \ref{tab:methods}, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
 
@@
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
-\jri{is this really chromosome 10 of maize?
-looks like only first 10Mb.
-if so, should mention that in figure legend}
 
@@
 \includegraphics[width=130mm]{figures/ThetasPlot.png}
-\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across chromosome 10 of {\it Zea mays} spp. {\it mays} using ANGSD-wrapper\label{fig:theta} }
+\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across the first 10 megabases of chromosome 10 in {\it Zea mays} spp. {\it mays} using ANGSD-wrapper\label{fig:theta} }
 \end{figure}

### 2016-01-13 manuscript.tex
add small results paragraph
results

@@
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, Patterson's D-statistic test of admixture \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.
+\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.\jri{are all these within ANGSD or are some ngsPopgen?}
 One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
@@
     \hline
-    Methods Implemented & Interactive Graphing \\ \hline
-    ABBA BABA & Yes \\ \hline
+    \textbf{Methods Implemented} & \textbf{Interactive Graphing} \\ \hline \hline
+    Patterson's D (ABBA-BABA) & Yes \\ \hline
     Admixture & Yes  \\ \hline
-    Ancestral Sequence & N/A \\ \hline
-    Genotype Likelihoods & N/A \\ \hline
-    Inbreeding Coefficients & N/A \\ \hline
+    Ancestral Sequence & No \\ \hline
+    Genotype Likelihoods & No \\ \hline
+    Inbreeding Coefficients & No \\ \hline
     PCA & Yes \\ \hline
     Site Frequency Spectrum & Yes \\ \hline
-    Theta Calculation & Yes \\ \hline
+    Diversity Statistics  & Yes \\ \hline
     \end{tabular}
@@
 \end{table}
+\jri{list diversity statistics in parentheses. what about Fst?}
 
@@
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
+\jri{add genes from gff.
+i don't think it's Mb 1-10, pls check.
+if so, is x-axis wrong?}
 
@@
 
-\section*{Results}
-
+\section*{Example data}
+\jri{show k=2 and k=3 for adxmxture in supp.}
 \begin{figure}
@@
 
-\section*{Discussion}
+As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
+We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10.\jri{cite Chia 2012 instead.
+include region \& mean depth, supp.
+table w/ sample names \& depth.
+include download link.}
+
+We first use \X\ to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \texit{Zea mays} ssp.
+\textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ since these are all highly inbred samples.
+Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication.\jri{cite Tim}
+We find further evidence of the effect of domestication using \X, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C).\jri{fill in \X's with which A-W script we used.
+probably need to modify this to mention Tripsacum and generation of the ancestral seq.
+and which A-W script we used for that.}
+Using \X, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
+The mean $F_{ST}$ in this region is \X, close to the genome-wide reported value of \X \citep{hufford2012comparative} .  
+The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative}
+
+We next include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
+Figure \ref{fig:admixture}) identifies previously unreported structure within the domesticated maize samples used in \citep{hufford2012comparative}
 
@@
 Finally, we would like to thank Felix Andrews for statistical advice, although we did not follow it.
-
+\jri{make sure figures/tables go where Mol. Ecol. Res. suggests (at end? in text?)}
+\clearpage
 \bibliographystyle{unsrt}

### 2016-01-13 manuscript.tex
more edits, comments

@@
 \includegraphics[width=\linewidth]{figures/mt4labeled.png}
-\caption{Admixture analysis for {\it Zea mays} spp. {\it mays}, {\it Zea mays} spp. {\it parviglumas}, and {\it Zea mays} spp. {mexicana} (Mex.) with K=4 source populations. }
+\caption{Admixture analysis for {\it Zea mays} spp.
+{\it mays}, {\it Zea mays} spp.
+{\it parviglumas}, and {\it Zea mays} spp.
+{mexicana} (\texit{mex}) with K=4 source populations.
+For a list of samples see Table \X.
+}
 \label{fig:admixture}
@@
 
-As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
+\jri{change Mex to mex and make italics in figure axis} As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
 We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10.\jri{cite Chia 2012 instead.
 include region \& mean depth, supp.
-table w/ sample names \& depth.
+table w/ sample names in order from fig.
+\& depth.
 include download link.}
 
-We first use \X\ to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \texit{Zea mays} ssp.
+We first use \X\ to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ since these are all highly inbred samples.
+\jri{do we worry that fig.
+3 not graphs directly from A-W?}
+
 Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication.\jri{cite Tim}
@@
 Using \X, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
-The mean $F_{ST}$ in this region is \X, close to the genome-wide reported value of \X \citep{hufford2012comparative} .  
-The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative}
+The mean $F_{ST}$ in this region is \X, close to the genome-wide reported value of \X \citep{hufford2012comparative}.\jri{change if not true}  
+The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff.
+make supp.
+table of genes in top 1\%.
+sorry to add, but this is cool \& demonstrates more stuff A-W can do.
+don't need a graph}
 
-We next include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
-Figure \ref{fig:admixture}) identifies previously unreported structure within the domesticated maize samples used in \citep{hufford2012comparative}
+Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
+Figure \ref{fig:admixture} identifies structure within the domesticated maize separating three high-latitude temperate landraces from the other tropical accessions. 
+We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis using SNP genotyping \jri{cite Hufford 2013 plos gen}.
+\textit{Zea mays} ssp.
+{mexicana} clusters into its own group, along with a single accession of ssp.
+{parviglumis} collected from region in which many teosinte populations appear to be the result of admixture between the two subspecies \jri{cite Fang 2012}.
+A single  ssp.
+{parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
 

### 2016-01-13 manuscript.tex
nothing to se ehere
addressed figure comments

@@
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.\jri{are all these within ANGSD or are some ngsPopgen?}
+\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.
+%\jri{are all these within ANGSD or are some ngsPopgen?} all are in ANGSD
 One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
@@
 \begin{center}
-    \begin{tabular}{ | l | p{5cm} |}
+    \begin{tabular}{ | p{5cm} | p{5cm} |}
     \hline
@@
     Site Frequency Spectrum & Yes \\ \hline
-    Diversity Statistics  & Yes \\ \hline
+    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
+    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
+    Fst & Yes \\ \hline
     \end{tabular}
@@
 \end{table}
-\jri{list diversity statistics in parentheses. what about Fst?}
+%\jri{in table: list diversity statistics in parentheses. what about Fst?} done. and added Neutrality test stats
 
@@
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
-\jri{add genes from gff.
+% \jri{add genes from gff.
 i don't think it's Mb 1-10, pls check.
-if so, is x-axis wrong?}
+if so, is x-axis wrong?} done
 
@@
 \centering
-\includegraphics[width=130mm]{figures/ThetasPlot.png}
-\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across the first 10 megabases of chromosome 10 in {\it Zea mays} spp. {\it mays} using ANGSD-wrapper\label{fig:theta} }
+\includegraphics[width=130mm]{figures/fig1.png}
+\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5 megabase region of chromosome 10 in {\it Zea mays} spp.
+{\it mays} using ANGSD-wrapper.
+Blue boxes indicate genic regions provided by a GFF annotation.
+\label{fig:theta} }
 \end{figure}
@@
 \section*{Example data}
-\jri{show k=2 and k=3 for adxmxture in supp.}
+% \jri{show k=2 and k=3 for adxmxture in supp.} done
 \begin{figure}
@@
 \includegraphics[width=\linewidth]{figures/figure3big.pdf}
-\caption{Summary statistics for {\it Zea mays}. Site frequency spectra for A. maize and B. teosinte. C. distribution of pairwise differences for maize (blue) and teosinte (orange).}
+\caption{Summary statistics for {\it Zea mays}.
+Site frequency spectra for A.
+maize and B.
+teosinte.
+C.
+distribution of pairwise differences for maize (blue) and teosinte (orange).
+Pairwise diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
 \label{fig:figure3}
@@
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ since these are all highly inbred samples.
-\jri{do we worry that fig.
-3 not graphs directly from A-W?}
+%\jri{do we worry that fig.
+3 not graphs directly from A-W?} % noted this in the figure legend
 

### 2016-01-13 manuscript.tex
added table S1

@@
 {mexicana} (\texit{mex}) with K=4 source populations.
-For a list of samples see Table \X.
+For a list of samples see Table \ref{tab:samples}.
 }
@@
 
-\jri{change Mex to mex and make italics in figure axis} As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
-We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10.\jri{cite Chia 2012 instead.
-include region \& mean depth, supp.
-table w/ sample names in order from fig.
-\& depth.
-include download link.}
+% \jri{change Mex to mex and make italics in figure axis} done
 
-We first use \X\ to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
+As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
+We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize} .\jri{include download link.}
+
+We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ since these are all highly inbred samples.
 %\jri{do we worry that fig.
-3 not graphs directly from A-W?} % noted this in the figure legend
+3 not graphs directly from A-W?} % noted this in the figure legend, SFSes are basically from AW, just colors changed.
+In the future should we add color picker for SFS?
 
-Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication.\jri{cite Tim}
-We find further evidence of the effect of domestication using \X, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C).\jri{fill in \X's with which A-W script we used.
+Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
+We find further evidence of the effect of domestication using the Thetas method, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C).
+These results use Tripsacum as an ancestral sequence, which was generated using the Ancestral Sequence method in ANGSD-wrapper.
+%\jri{fill in \X's with which A-W script we used.
 probably need to modify this to mention Tripsacum and generation of the ancestral seq.
-and which A-W script we used for that.}
-Using \X, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
+and which A-W script we used for that.} done
+Using the 2DSFS method, which includes an $F_{ST}$ calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
 The mean $F_{ST}$ in this region is \X, close to the genome-wide reported value of \X \citep{hufford2012comparative}.\jri{change if not true}  
@@
 Figure \ref{fig:admixture} identifies structure within the domesticated maize separating three high-latitude temperate landraces from the other tropical accessions. 
-We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis using SNP genotyping \jri{cite Hufford 2013 plos gen}.
+We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis using SNP genotyping \citep{hufford2013genomic}.
 \textit{Zea mays} ssp.
 {mexicana} clusters into its own group, along with a single accession of ssp.
-{parviglumis} collected from region in which many teosinte populations appear to be the result of admixture between the two subspecies \jri{cite Fang 2012}.
+{parviglumis} collected from region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.
 A single  ssp.

### 2016-01-14 manuscript.tex
added mean fst

@@
 Using the 2DSFS method, which includes an $F_{ST}$ calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
-The mean $F_{ST}$ in this region is \X, close to the genome-wide reported value of \X \citep{hufford2012comparative}.\jri{change if not true}  
+The mean $F_{ST}$ in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
 The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff.

### 2016-01-14 manuscript.tex
added data link

@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
-We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize} .\jri{include download link.}
+We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize} .
+The data is available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.
 

### 2016-01-15 manuscript.tex
table, fst

@@
 \newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
-
+\newcommand{\fst}{$\textrm{F_{\textrm{\scriptsize ST}}}$}          
 \newcommand{\X}{\textcolor{red}{\bf X}}
 
-\title{More than one Author with different Affiliations}
+%\title{More than one Author with different Affiliations}
 \author[1,$\dag$]{Arun Durvasula}
@@
 \begin{center}
-    \begin{tabular}{ | p{5cm} | p{5cm} |}
+    \begin{tabular}{ | p{3.5cm} | p{5cm} | p{2.5cm}  |}
     \hline
-    \textbf{Methods Implemented} & \textbf{Interactive Graphing} \\ \hline \hline
-    Patterson's D (ABBA-BABA) & Yes \\ \hline
-    Admixture & Yes  \\ \hline
-    Ancestral Sequence & No \\ \hline
-    Genotype Likelihoods & No \\ \hline
-    Inbreeding Coefficients & No \\ \hline
-    PCA & Yes \\ \hline
-    Site Frequency Spectrum & Yes \\ \hline
-    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
-    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
-    Fst & Yes \\ \hline
+    \textbf{Method} & \textbf{Calculations} & \textbf{Interactive Graphing} \\ \hline \hline
+    SFS     &    Site frequency spectrum  & Yes  \\ \hline     2DSFS   &    Joint site frequency spectrum, \fst & Yes  \\ \hline 
+    Abbababa  &  Patterson's D statistic & Yes  \\ \hline 
+    Ancestral &  Extract ancestral sequence from BAM file  & No \\ \hline 
+    Genotypes &  Genotype likelihood estimations & No  \\ \hline 
+    PCA     &    Principal component analysis & Yes  \\ \hline 
+    Thetas   &   Diversity statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$) and neutrality tests (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes  \\ \hline
+    Inbreeding & Calculate per-individual inbreeding coefficients with ngsF & No  \\ \hline 
+    Admixture  & Perform admixture analysis & Yes  \\ \hline
+
+ %   Patterson's D (ABBA-BABA) & Yes \\ \hline
+%    Admixture & Yes  \\ \hline
+%    Ancestral Sequence & No \\ \hline
+%    Genotype Likelihoods & No \\ \hline
+%    Inbreeding Coefficients & No \\ \hline
+%    PCA & Yes \\ \hline
+%    Site Frequency Spectrum & Yes \\ \hline
+%    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
+%    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
+%    $\textrm{F_{\textrm{\scriptsize ST}}}$ & Yes \\ \hline
     \end{tabular}
@@
 \end{table}
-%\jri{in table: list diversity statistics in parentheses. what about Fst?} done. and added Neutrality test stats
 
@@
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
-% \jri{add genes from gff.
-i don't think it's Mb 1-10, pls check.
-if so, is x-axis wrong?} done
 
@@
 \section*{Example data}
-% \jri{show k=2 and k=3 for adxmxture in supp.} done
 \begin{figure}
@@
 
-% \jri{change Mex to mex and make italics in figure axis} done
-
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
@@
 and which A-W script we used for that.} done
-Using the 2DSFS method, which includes an $F_{ST}$ calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
-The mean $F_{ST}$ in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
-The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff.
+Using the 2DSFS method, which includes an \fst calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
+The mean \fst in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
+The \X windows in the 1\% tail of highest \fst values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff.
 make supp.

### 2016-01-15 manuscript.tex
edits

@@
 \newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
-\newcommand{\fst}{$\textrm{F_{\textrm{\scriptsize ST}}}$}          
+\newcommand{\fst}{${F_{ST}}$ }          
 \newcommand{\X}{\textcolor{red}{\bf X}}
@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
-We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize} .
+We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
 The data is available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.
+In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.
 
 We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
-\textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ since these are all highly inbred samples.
-%\jri{do we worry that fig.
-3 not graphs directly from A-W?} % noted this in the figure legend, SFSes are basically from AW, just colors changed.
-In the future should we add color picker for SFS?
-
+\textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
 Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
-We find further evidence of the effect of domestication using the Thetas method, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C).
-These results use Tripsacum as an ancestral sequence, which was generated using the Ancestral Sequence method in ANGSD-wrapper.
-%\jri{fill in \X's with which A-W script we used.
-probably need to modify this to mention Tripsacum and generation of the ancestral seq.
-and which A-W script we used for that.} done
+We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
+These results use a sample from the genus \textit{Tripsacum} as an ancestral sequence generated using the Ancestral Sequence method. 
 Using the 2DSFS method, which includes an \fst calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
 The mean \fst in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
-The \X windows in the 1\% tail of highest \fst values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff.
-make supp.
-table of genes in top 1\%.
-sorry to add, but this is cool \& demonstrates more stuff A-W can do.
-don't need a graph}
+The \X windows in the 1\% tail of highest \fst values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative}
 

### 2016-01-17 manuscript.tex
changed to PDFs
draft done

@@
 \citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.
-%\jri{are all these within ANGSD or are some ngsPopgen?} all are in ANGSD
 One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
@@
 \centering
-\includegraphics[width=130mm]{figures/fig1.png}
+\includegraphics[width=130mm]{figures/fig1.pdf}
 \caption{A visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5 megabase region of chromosome 10 in {\it Zea mays} spp.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/mt4labeled.png}
+\includegraphics[width=\linewidth]{figures/mt4labeled.pdf}
 \caption{Admixture analysis for {\it Zea mays} spp.
@@
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
+The SFS and diversity statistics were calculated using ancestral states inferred in the Ancestral Sequence method from a single resequenced genome of \textit{Tripsacum}.  
 Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
 We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
-These results use a sample from the genus \textit{Tripsacum} as an ancestral sequence generated using the Ancestral Sequence method. 
-Using the 2DSFS method, which includes an \fst calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
-The mean \fst in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
-The \X windows in the 1\% tail of highest \fst values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative}
+Using the 2DSFS method, which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \citet{hufford2012comparative}.
+There are no loci in this region that have previously been identified as showing evidence of selection during domestication, and these data also do not find any large consecutive regions of high \fst.
 
@@
 Finally, we would like to thank Felix Andrews for statistical advice, although we did not follow it.
-\jri{make sure figures/tables go where Mol. Ecol. Res. suggests (at end? in text?)}
 \clearpage

### 2016-01-17 manuscript.tex
manuscript edits

@@
 Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
-ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses and provides interactive graphing of ANGSD results to enhance data exploration.
+ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
+ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
 We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
@@
 
-High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or survey fine-scale patterns of variation along the genome.  
-Application of these methods has led to changes in both experimental design and data analysis \citep{ekblom2011applications}.
-Many of the popular software packages used by researchers \citep[see][]{excoffier2006computer}  were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated.
+High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or survey fine-scale patterns of variation across the genome.  
+Application of these approaches has led to changes in both experimental design and data analysis \citep{ekblom2011applications}.
+Many popular software packages used by researchers for analysis of comparative resequencing data \citep[see][]{excoffier2006computer} were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated.
 In particular, short read sequencing has brought new challenges, including highly variable coverage, missing data, and high per-nucleotide error rates.
 
-A number of tools have recently been published to handle high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}, but the majority of these either make limiting assumptions about the data (e.g., all sites have been sequenced, all genomes are haploid, sequencing is to sufficient depth, all individuals are outcrossing) or are specialized tools offering a narrow set of analysis options.
+A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}, but the majority of these either make limiting assumptions about the data (e.g., all sites have been sequenced, all genomes are haploid, sequencing is to sufficient depth, all individuals are outcrossing) or are specialized tools offering a narrow set of analysis options.
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and neutrality test statistics \citep{korneliussen_calculation_2013}.
+\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
 One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
@@
 Here we present ANGSD-wrapper, a user-friendly interface to ANGSD. 
-ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines inherent in ANGSD as well as pipelines involving related programs such as ngsPopGen, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
-Because the large volume of data associated with high throughput sequence analysis is often difficult to explore by hand, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
-We demonstrate some of the analyses possible using ANGSD-wrapper using low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
+ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD.
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
+Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
+We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
 ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
@@
 \section*{Methods}
-ANGSD-wrapper is a set of configuration files and scripts written in the Bash UNIX shell.  
+ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash UNIX shell.  
 The scripts can be run either on a standalone computer with a UNIX terminal, or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
@@
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
-ANGSD-wrapper provides a common configuration file, common.conf, which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
+ANGSD-wrapper provides a common configuration file, "common.conf," which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
 In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
@@
 
-Analyses supported by ANGSD-wrapper are shown in table \ref{tab:methods}, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
+Analyses supported by ANGSD-wrapper are shown in Table \ref{tab:methods}, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
 
-The visualization software included with ANGSD-wrapper is contained within it's own directory called shinyGraphing.
+The visualization software included with ANGSD-wrapper is contained within it's own directory called "shinyGraphing."
 This application must be started in R and can be accessed locally from a web browser.
@@
 
-In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate file of results. 
+In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate results file. 
 The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
-We used resequenced samples from the HapMap2 project  and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
+We used resequenced samples from the Maize HapMap2 project and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
 The data is available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.
@@
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
-The SFS and diversity statistics were calculated using ancestral states inferred in the Ancestral Sequence method from a single resequenced genome of \textit{Tripsacum}.  
+The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum}.  
 Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
@@
 \section*{Conclusions}
-Our software ANGSD-wrapper provides an intuitive and easy-to-use interface to employ the powerful and flexible suite of population genetic analyses developed in ANGSD \citep{korneliussen2014angsd} and permits the exploration of genome-scale results through interactive visualization.
+ANGSD-wrapper provides an easy-to-use interface that simplifies many population genetic analyses implemented in ANGSD \citep{korneliussen2014angsd} and permits the exploration of genome-scale results through interactive visualization.
 ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.
@@
 \section*{Acknowledgements}
-We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM) and funding from the UC Davis Plant Sciences Department.
-We thank members of the Ross-Ibarra and Morrell labs for discussion and software testing.
+We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM), from a University of Minnesota Doctoral Dissertation Fellowship supporting TJYK, and from the University of California Davis Plant Sciences Department.
+We thank members of the Ross-Ibarra and Morrell Labs for discussion and software testing.
 We thank the authors of ANGSD and related programs for answering questions, particularly Matteo Fumagalli and Filipe Vieira.

### 2016-01-17 manuscript.tex
edits

@@
 \newcommand{\fst}{${F_{ST}}$ }          
-\newcommand{\X}{\textcolor{red}{\bf X}}
+\newcommand{\X}{\textcolor{red}{\bf X }}
 
@@
 ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen\jri{please add citation to url for git repo}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
@@
     SFS     &    Site frequency spectrum  & Yes  \\ \hline     2DSFS   &    Joint site frequency spectrum, \fst & Yes  \\ \hline 
-    Abbababa  &  Patterson's D statistic & Yes  \\ \hline 
+    ABBA/BABA  &  Patterson's D statistic & Yes  \\ \hline 
     Ancestral &  Extract ancestral sequence from BAM file  & No \\ \hline 
@@
 
-Analyses supported by ANGSD-wrapper are shown in Table \ref{tab:methods}, and a detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
+Analyses supported by ANGSD-wrapper are shown in Table \ref{tab:methods}.
+A detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
 
@@
 \centering
-\includegraphics[width=130mm]{figures/fig1.pdf}
+\includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
 \caption{A visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5 megabase region of chromosome 10 in {\it Zea mays} spp.
@@
 
-\section*{Example data}
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
-\caption{Example analysis workflow diagram.}
-\label{fig:workflow}
-\end{figure}
+%\section*{Example data}
+%\begin{figure}
+%\centering
+%\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
+%\caption{Example analysis workflow diagram.}
+%\label{fig:workflow}
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/figure3big.pdf}
+\includegraphics[width=0.8\linewidth]{figures/figure3big.pdf}
 \caption{Summary statistics for {\it Zea mays}.
@@
 
+As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
+We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) \jri{please add Tripsacum to this table and add species identity to the table} and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
+The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.
+In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.
+
+We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
+\textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
+The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
+Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of \X and \X in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
+We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
+Using the 2DSFS method, which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
+There are no loci in this region that have previously been identified as showing evidence of selection during domestication, and these data also do not find any large consecutive regions of high \fst.
+
 \begin{figure}
 \centering
-\includegraphics[width=\linewidth]{figures/mt4labeled.pdf}
+\includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
 \caption{Admixture analysis for {\it Zea mays} spp.
@@
 {mexicana} (\texit{mex}) with K=4 source populations.
-For a list of samples see Table \ref{tab:samples}.
+Individuals are shown in same order as Table \ref{tab:samples}.
 }
@@
 
-As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of a subset of domesticated maize and wild teosinte.
-We used resequenced samples from the Maize HapMap2 project and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
-The data is available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.
-In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.
-
-We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
-\textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
-The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum}.  
-Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
-We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
-Using the 2DSFS method, which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \citet{hufford2012comparative}.
-There are no loci in this region that have previously been identified as showing evidence of selection during domestication, and these data also do not find any large consecutive regions of high \fst.
-
 Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
-Figure \ref{fig:admixture} identifies structure within the domesticated maize separating three high-latitude temperate landraces from the other tropical accessions. 
-We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis using SNP genotyping \citep{hufford2013genomic}.
+Using the Admixture method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).\jri{parviglumis spelled wrong in figure legend}
+We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
 \textit{Zea mays} ssp.

### 2016-01-19 manuscript.tex
addressed ngspopgen citation

@@
 ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen\jri{please add citation to url for git repo}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.

### 2016-01-20 manuscript.tex
added species to table
added TajD and pi

@@
 \newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
+\newcommand{\ad}[1]{\todo[size=\scriptsize, color=blue]{#1}}
 \newcommand{\fst}{${F_{ST}}$ }          
@@
 The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of \X and \X in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
+Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X (Maize=0.0061, Teo=0.0082)\ad{wasn't sure if you wanted maize pi only or both so I included both}, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
 We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
@@
 {\it mays}, {\it Zea mays} spp.
-{\it parviglumas}, and {\it Zea mays} spp.
+{\it parviglumis}, and {\it Zea mays} spp.
 {mexicana} (\texit{mex}) with K=4 source populations.

### 2016-01-20 manuscript.tex
fix parv spelling

@@
 Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
-Using the Admixture method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).\jri{parviglumis spelled wrong in figure legend}
+Using the Admixture method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
 We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.

### 2016-01-20 manuscript.tex
added trip coverage

@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
-We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) \jri{please add Tripsacum to this table and add species identity to the table} and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
+We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
 The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}.

### 2016-01-20 manuscript.tex
fixed pi in text

@@
 The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X (Maize=0.0061, Teo=0.0082)\ad{wasn't sure if you wanted maize pi only or both so I included both}, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
+We find mean levels of nucleotide diversity in this region to be 0.0061 and 0.0082 (maize and teosinte, respectively), which is consistent with previous genome wide results (0.0072 in maize and 0.0078 in teosinte \citep{hufford2012comparative}).
+We also show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
 We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 

### 2016-01-25 manuscript.tex
for_submission

--- 2016-01-20 manuscript.tex
@@
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
-ANGSD-wrapper provides a common configuration file, "common.conf," which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
+ANGSD-wrapper provides a common configuration file, ``common.conf,'' which  holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
 In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
@@
 
-The visualization software included with ANGSD-wrapper is contained within it's own directory called "shinyGraphing."
+The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
 This application must be started in R and can be accessed locally from a web browser. 
@@
 \includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
-\caption{A visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5 megabase region of chromosome 10 in {\it Zea mays} spp.
+\caption{Visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5Mb region of chromosome 10 in {\it Zea mays} ssp.
 {\it mays} using ANGSD-wrapper.
@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
-We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10 megabase region on chromosome 10 \citep{chia2012maize}.
+We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
 The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}. 
@@
 
-We first use the SFS method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
+We first use the ``SFS'' method to estimate the site frequency spectrum (SFS) of both maize and its wild progenitor \textit{Zea mays} ssp.
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
-The SFS and diversity statistics were calculated using ancestral states inferred in the "Ancestral Sequence" method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-Consistent with previous results \citep{hufford2012comparative}, we find mean levels of nucleotide diversity in this region to be \X (Maize=0.0061, Teo=0.0082)\ad{wasn't sure if you wanted maize pi only or both so I included both}, and show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
-We find further evidence of the effect of domestication using the Thetas method, revealing lower overall levels of diversity in maize (Figure \ref{fig:figure3}C). 
-Using the 2DSFS method, which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
-There are no loci in this region that have previously been identified as showing evidence of selection during domestication, and these data also do not find any large consecutive regions of high \fst.
+The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
+We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.\jri{cite Eyre-Walker 1998 PNAS here instead}
+Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with  mean levels of nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
+Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
+There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis.
 
@@
 \includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
-\caption{Admixture analysis for {\it Zea mays} spp.
-{\it mays}, {\it Zea mays} spp.
-{\it parviglumis}, and {\it Zea mays} spp.
-{mexicana} (\texit{mex}) with K=4 source populations.
-Individuals are shown in same order as Table \ref{tab:samples}.
-}
+\caption{Admixture analysis for {\it Zea mays} ssp. {\it mays}, {\it Zea mays} ssp. {\it parviglumis}, and {\it Zea mays} ssp. {mexicana} (\textit{mex}) with K=4 source populations. 
+Individuals are shown in the same order as Table \ref{tab:samples}. }
 \label{fig:admixture}
@@
 
-Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. {mexicana} to assess evidence for admixture.  
-Using the Admixture method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
-We find no evidence of admixture between these lowland maize samples and  ssp. {mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
+Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. \textit{mexicana} to assess evidence for admixture.  
+Using the ``Admixture'' method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify  structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
+We find no evidence of admixture between these lowland maize samples and  ssp. \textit{mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
 \textit{Zea mays} ssp.
-{mexicana} clusters into its own group, along with a single accession of ssp.
-{parviglumis} collected from region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.
+\textit{mexicana} clusters into its own group, along with a single accession of ssp.
+\textit{parviglumis} collected from a region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.
 A single  ssp.
-{parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
+\textit{parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
 

### 2016-01-26 manuscript.tex
almost...
addressed Fst and citation

@@
 The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.\jri{cite Eyre-Walker 1998 PNAS here instead}
+We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation} \citep{Beissinger031666}.
 Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with  mean levels of nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
-There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis.
+There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis (Figure \ref{fig:suppfst}).
 

### 2016-01-26 manuscript.tex
italics
TJYK edits to intro

@@
 Many popular software packages used by researchers for analysis of comparative resequencing data \citep[see][]{excoffier2006computer} were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated.
-In particular, short read sequencing has brought new challenges, including highly variable coverage, missing data, and high per-nucleotide error rates.
+Despite the decreasing cost of sequencing, researchers must allocate finite resources and balance the depth of sequencing and the breadth of a sample. 
+This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample. 
+This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition. 
+%TJYK: CITE Pluzhnikov and Donnelly 1996 and Felsenstein 2005
+While the experimental design challenges with moleclar population genetic studies have existed for at least a decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
 
-A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}, but the majority of these either make limiting assumptions about the data (e.g., all sites have been sequenced, all genomes are haploid, sequencing is to sufficient depth, all individuals are outcrossing) or are specialized tools offering a narrow set of analysis options.
+A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}.
+However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
-One of the most important features of ANGSD is that most analyses are performed directly on genotype likelihoods, freeing users from the requirement of calling variants or genotypes and permitting analysis of low-coverage data or sequences with large amounts of missing data.
+\citep{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
+ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
+One of the most important features of ANGSD is that analyses are integreated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
+This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage.
 
-Here we present ANGSD-wrapper, a user-friendly interface to ANGSD. 
+Here we present ANGSD-wrapper, a user-friendly interface to stable versions of ANGSD. 
 ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
 ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
+Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.

### 2016-01-27 manuscript.tex
addressed citation issues
added more about admixture

@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample. 
-This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition. 
-%TJYK: CITE Pluzhnikov and Donnelly 1996 and Felsenstein 2005
-While the experimental design challenges with moleclar population genetic studies have existed for at least a decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
+This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition \citep{pluzhnikov1996optimal, felsenstein2005using}. 
+While the experimental design challenges with molecular population genetic studies have existed for at least a decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
 
@@
 ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
-One of the most important features of ANGSD is that analyses are integreated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
+One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage.
@@
     Admixture  & Perform admixture analysis & Yes  \\ \hline 
-
- %   Patterson's D (ABBA-BABA) & Yes \\ \hline
-%    Admixture & Yes  \\ \hline
-%    Ancestral Sequence & No \\ \hline
-%    Genotype Likelihoods & No \\ \hline
-%    Inbreeding Coefficients & No \\ \hline
-%    PCA & Yes \\ \hline
-%    Site Frequency Spectrum & Yes \\ \hline
-%    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
-%    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
-%    $\textrm{F_{\textrm{\scriptsize ST}}}$ & Yes \\ \hline
     \end{tabular}
@@
 
-%\section*{Example data}
-%\begin{figure}
-%\centering
-%\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
-%\caption{Example analysis workflow diagram.}
-%\label{fig:workflow}
-%\end{figure}
-
 \begin{figure}
@@
 The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation} \citep{Beissinger031666}.
+We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
 Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
@@
 \includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
-\caption{Admixture analysis for {\it Zea mays} ssp. {\it mays}, {\it Zea mays} ssp. {\it parviglumis}, and {\it Zea mays} ssp. {mexicana} (\textit{mex}) with K=4 source populations. 
+\caption{Admixture analysis for {\it Zea mays} ssp. {\it mays}, {\it Zea mays} ssp. {\it parviglumis}, and {\it Zea mays} ssp. {\it mexicana} (\textit{mex}) with K=4 source populations. 
 Individuals are shown in the same order as Table \ref{tab:samples}. }
@@
 Using the ``Admixture'' method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
-We find no evidence of admixture between these lowland maize samples and  ssp. \textit{mexicana}, consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
+This analysis is a supervised method that assigns proportions of the each individual's genome to a putative source population, represented by the colors. 
+In maize, we see that most individuals come from one source population (dark grey or dark blue), which is evidence for no admixture between these lowland maize samples and  ssp.
+\textit{mexicana}, and is consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
 \textit{Zea mays} ssp.
-\textit{mexicana} clusters into its own group, along with a single accession of ssp.
+\textit{mexicana} clusters into its own group (orange), along with a single accession of ssp.
 \textit{parviglumis} collected from a region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.
@@
 \clearpage
-\bibliographystyle{unsrt}
+\bibliographystyle{apa}
 \bibliography{references}

### 2016-01-27 manuscript.tex
more analysis
edits

@@
 {\it mays} using ANGSD-wrapper.
-Blue boxes indicate genic regions provided by a GFF annotation.
+Darker colors indicate a higher density of points.
+Blue boxes indicate gene annotations provided by a GFF file.
 \label{fig:theta} }
@@
 \caption{Summary statistics for {\it Zea mays}.
-Site frequency spectra for A.
+Derived site frequency spectra for A.
 maize and B.
@@
 \end{figure}
-
+\jri{figure says "spectrum" should be plural}
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
@@
 We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
-Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
+Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
@@
 \includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
-\caption{Admixture analysis for {\it Zea mays} ssp. {\it mays}, {\it Zea mays} ssp. {\it parviglumis}, and {\it Zea mays} ssp. {\it mexicana} (\textit{mex}) with K=4 source populations. 
+\caption{Admixture analysis for {\it Zea mays} ssp.
+{\it mays}, {\it Zea mays} ssp.
+{\it parviglumis}, and {\it Zea mays} ssp.
+{\it mexicana} (\textit{mex}) with colors representing the K=4 source populations.
 Individuals are shown in the same order as Table \ref{tab:samples}. }
@@
 \end{figure}
-
+\jri{what are vertical lines in pi density plot?
+need to be consistent about saying "nucleotide diversity" vs "pairwise diversity" vs "pairwise differences" etc.
+I vote for "pairwise nucleotide diversity"}
 Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. \textit{mexicana} to assess evidence for admixture.  
@@
 This analysis is a supervised method that assigns proportions of the each individual's genome to a putative source population, represented by the colors. 
+We ran this analysis over a range of number of ancestral populations (K= 2, 3, 4; See Figure \ref{fig:suppadmix}) and chose the value with the maximum likelihood while keeping the number of parameters as low as possible.
 In maize, we see that most individuals come from one source population (dark grey or dark blue), which is evidence for no admixture between these lowland maize samples and  ssp.
@@
 ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.  
-
+\jri{why is table in middle of refs? check refs for completeness. e.g. skotte is missing page numbers}
 \section*{Acknowledgements}

### 2016-01-27 manuscript.tex
spectra

@@
 \end{figure}
-\jri{figure says "spectrum" should be plural}
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 

### 2016-01-28 manuscript.tex
fixed felsenstein

@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample. 
-This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition \citep{pluzhnikov1996optimal, felsenstein2005using}. 
+This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
 While the experimental design challenges with molecular population genetic studies have existed for at least a decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.

### 2016-01-28 manuscript.tex
updates

@@
 \newcommand{\ad}[1]{\todo[size=\scriptsize, color=blue]{#1}}
+\newcommand{\ph}[1]{\todo[size=\scriptsize, color=pink]{#1}}
 \newcommand{\fst}{${F_{ST}}$ }          
@@
 Korneliussen et al.
-\citep{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
+\ph{Should be just \(2014\)} \citep{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
 ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
@@
 \section*{Methods}
-ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash UNIX shell.  
-The scripts can be run either on a standalone computer with a UNIX terminal, or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
+\ph{change of language} ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language.%UNIX shell.  
+The scripts can be run either on a standalone computer with a UNIX terminal, \ph{has anyone tested AW on a cluster?} or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
+A Python installation \ph{Citation} is required for some light, dynamic pre-processing of the data.
 An installation of the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.  
-The visualization portion of ANGSD-wrapper also requires installation of the R packages shiny \citep{shiny}, genomeIntervals \citep{genomeIntervals}, and ape \citep{APE}.
+The visualization portion of ANGSD-wrapper also requires installation of the R packages shiny \citep{shiny}, genomeIntervals \citep{genomeIntervals}, ape \citep{APE}, lattice \ph{citation}, Hmisc \ph{citation}, data.table \ph{citation}, DT \ph{citation}, and shinythemes \ph{citation}.
 
@@
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
-ANGSD-wrapper provides a common configuration file, ``common.conf,'' which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
+ANGSD-wrapper provides a common configuration file, \ph{renamed} \texttt{Common\_Config}, which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
 In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
@@
 The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
-This application must be started in R and can be accessed locally from a web browser. 
+This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope
 This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  

### 2016-01-29 manuscript.tex
More notes, updates, and citations

@@
 Here we present ANGSD-wrapper, a user-friendly interface to stable versions of ANGSD. 
-ANGSD-wrapper takes the form of a set of configuration files and 'wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
+ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
 ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
@@
 \section*{Methods}
-\ph{change of language} ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language.%UNIX shell.  
+\ph{change of language} ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language. %UNIX shell.  
 The scripts can be run either on a standalone computer with a UNIX terminal, \ph{has anyone tested AW on a cluster?} or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
-A Python installation \ph{Citation} is required for some light, dynamic pre-processing of the data.
+A Python installation \citep{Python} is required for some light, dynamic pre-processing of the data.
 An installation of the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.  
-The visualization portion of ANGSD-wrapper also requires installation of the R packages shiny \citep{shiny}, genomeIntervals \citep{genomeIntervals}, ape \citep{APE}, lattice \ph{citation}, Hmisc \ph{citation}, data.table \ph{citation}, DT \ph{citation}, and shinythemes \ph{citation}.
+The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor} for the visualization package.
+These dependent packages are installed automatically upon first run of visualization interface.
 
@@
     \hline
-    \textbf{Method} & \textbf{Calculations} & \textbf{Interactive Graphing} \\ \hline \hline
+    \textbf{Method} & \textbf{Calculations} & \textbf{Interactive Graphing} \\ \hline
     SFS     &    Site frequency spectrum  & Yes  \\ \hline     2DSFS   &    Joint site frequency spectrum, \fst & Yes  \\ \hline 
@@
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
-ANGSD-wrapper provides a common configuration file, \ph{renamed} \texttt{Common\_Config}, which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
+ANGSD-wrapper provides a common configuration file, \texttt{Common\_Config}, which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
 In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
@@
 ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.  
-\jri{why is table in middle of refs? check refs for completeness. e.g. skotte is missing page numbers}
 \section*{Acknowledgements}

### 2016-01-29 manuscript.tex
Cleanup of intro, add comment about results section

@@
 \newcommand{\ph}[1]{\todo[size=\scriptsize, color=pink]{#1}}
+\newcommand{\tjyk}[1]{\todo[size=\scriptsize, color=green]{#1}}
 \newcommand{\fst}{${F_{ST}}$ }          
@@
 A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}.
-However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
+However, few of these tools offer the flexibility to handle data with low coverage and high missingness, resulting in either high data loss or bias in estimates. 
 Korneliussen et al.
@@
 ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as \tjyk{Fumagali citation doesn't show up properly}ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
 Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
@@
 
-The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
+The visualization software included with ANGSD-wrapper is contained within its own directory called ``shinyGraphing.''
 This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope
@@
 \end{figure}
+\tjyk{Should this start a "Results" section?}
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 

### 2016-01-29 manuscript.tex
Minor changes to citations

--- 2016-01-29 manuscript.tex
@@
 However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
-Korneliussen et al.
-\ph{Should be just \(2014\)} \citep{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
+\citet{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
 ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
@@
 ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
+\ph{This could be reworded} Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
 Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
@@
 
-The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
+\ph{Necessary?}The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
 This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope

### 2016-01-29 manuscript.tex

@@
 A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}.
-However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
  \citet{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
@@
 ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as \tjyk{Fumagali citation doesn't show up properly}ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
 \ph{This could be reworded} Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.

### 2016-01-29 manuscript.tex
Changed some stuff stylistically

@@
 ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as \tjyk{Fumagali citation doesn't show up properly}ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as \tjyk{Fumagali citation doesn't show up properly}\ph{I think the citation is just 'Fumagali' and the link to the webpage.
+If so, then it works for me -PH}ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
 \ph{This could be reworded} Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
-Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
+Additionally, the wrapper scripts are written against `frozen' versions of ANGSD and supporting tools, for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
@@
 \section*{Methods}
-\ph{change of language} ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language. %UNIX shell.  
+ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language. 
 The scripts can be run either on a standalone computer with a UNIX terminal, \ph{has anyone tested AW on a cluster?} or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.

### 2016-01-29 cover_letter.tex
cl updates

@@
-ANGSD-wrapper is intended to simplify analysis and visualization of next generation resequencing data.
+Dear Managing Editor,
 
-Visualization of analysis results is implemented using Shiny graphics in R.
+Enclosed you will find our manuscript, "ANGSD-wrapper: utilities for analyzing next generation sequencing data."
+This manuscript reports the novel software, ANGSD-wrapper, which is intended to simplify analysis and visualization of next generation resequencing data.
+The focus of this program is on usability and it extends the previously published software, ANGSD and programs related to it. 
+ANGSD-wrapper consists of two parts: 1) ANGSD pipelines that simplify multi-step analyses into a single step, and 2) interactive graphing based on the direct output of ANGSD-wrapper. 
+Interactive visualization of analysis results is implemented using Shiny graphics in R, which uses modern web technologies such as Javascript and HTML5. 
+The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.
+We demonstrate the usefulness of the software by analyzing previously published data in maize and teosinte.
 
-The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.
+We believe that usability and usefulness of our software will appeal to the readership of Molecular Ecology Resources.
+
+Thank you,
+
+Arun Durvasula and Jeffrey Ross-Ibarra

### 2016-01-29 manuscript.tex

@@
 \definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
 %\setlength{\marginparwidth}{1.5cm}
@@
 \newcommand{\tjyk}[1]{\todo[size=\scriptsize, color=green]{#1}}
+\newcommand{\cl}[1]{\todo[size=\scriptsize, color=amethyst]{#1}}
 \newcommand{\fst}{${F_{ST}}$ }          
@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
+\cl{added text}
+ANGSD uses probabilistic approaches to calculate genome wide descriptive statistics.
 The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
 Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
+\cl{added text}
+This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
 ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.

### 2016-01-29 manuscript.tex
addressed comments

--- 2016-01-29 manuscript.tex
@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
+% \cl{added text}
+ANGSD uses probabilistic approaches to calculate genome wide descriptive statistics.
 The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
 Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
+% \cl{added text}
+This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
 ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
@@
 ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as \tjyk{Fumagali citation doesn't show up properly}\ph{I think the citation is just 'Fumagali' and the link to the webpage.
-If so, then it works for me -PH}ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
-\ph{This could be reworded} Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
+%\tjyk{Fumagali citation doesn't show up properly}\ph{I think the citation is just 'Fumagali' and the link to the webpage.
+If so, then it works for me -PH} \ad{yes, there's no citation other than the github repo (and no DOI there)}
+% \ph{This could be reworded} 
+Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
 Additionally, the wrapper scripts are written against `frozen' versions of ANGSD and supporting tools, for consistency of analysis. 
@@
 ANGSD-wrapper is a set of configuration files and scripts written primarily in the Bash scripting language. 
-The scripts can be run either on a standalone computer with a UNIX terminal, \ph{has anyone tested AW on a cluster?} or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
+The scripts can be run either on a standalone computer with a UNIX terminal, or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
+% \ph{has anyone tested AW on a cluster?} \ad{yep I have.
+I logged into a compute node via srun and ran it that way}
 A Python installation \citep{Python} is required for some light, dynamic pre-processing of the data.
@@
 
-\ph{Necessary?}The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
-This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope
+% \ph{Necessary?} \ad{changed the wording}
+The visualization software included with ANGSD-wrapper is contained in a directory called ``shinyGraphing.''
+This application is started from within ANGSD-wrapper and launched locally from a standard web browser. 
 This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
@@
 
+\section*{Results}
+
 \begin{figure}
@@
 C.
-distribution of pairwise differences for maize (blue) and teosinte (orange).
-Pairwise diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
+distribution of pairwise nucleotide diversity for maize (blue) and teosinte (orange).
+Pairwise nucleotide diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
 \label{fig:figure3}
 \end{figure}
-\tjyk{Should this start a "Results" section?}
-As a demonstration of analyses in ANGSD-wrapper, we explore patterns of diversity in a single genomic region of domesticated maize and wild teosinte. 
+
+%\tjyk{Should this start a "Results" section?} \ad{yes I think so.}
+
+As a demonstration of analyses in ANGSD-wrapper, we explore patterns of pairwise nucleotide diversity in a single genomic region of domesticated maize and wild teosinte. 
 We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
@@
 We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
-Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
+Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively).
+Figure \ref{fig:figure3}C shows the distribution of pairwise nucleotide diversity in each window within our region of interest and each vertical line represents the number of windows with the corresponding value of $\pi$.
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
@@
 \end{figure}
-\jri{what are vertical lines in pi density plot?
+% \jri{what are vertical lines in pi density plot?
 need to be consistent about saying "nucleotide diversity" vs "pairwise diversity" vs "pairwise differences" etc.
-I vote for "pairwise nucleotide diversity"}
+I vote for "pairwise nucleotide diversity"} \ad{changed.}
 Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. \textit{mexicana} to assess evidence for admixture.  

### 2016-01-29 manuscript.tex
Made additional edits to abstract
ill comment later
edit

@@
 \newcommand{\cl}[1]{\todo[size=\scriptsize, color=amethyst]{#1}}
+\newcommand{\tvk}[1]{\todo[size=\scriptsize, color=razzledazzlerose]{#1}}
 \newcommand{\fst}{${F_{ST}}$ }          
@@
 This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
-ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
+%ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
 ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
@@
 Figure \ref{fig:figure3}C shows the distribution of pairwise nucleotide diversity in each window within our region of interest and each vertical line represents the number of windows with the corresponding value of $\pi$.
+\jri{no dude, not explaining what a histogram is, explain what the vertical line is (mean, median, etc.)}
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.

### 2016-01-29 manuscript.tex
oh thats what jeff meant

--- 2016-01-29 manuscript.tex
@@
 distribution of pairwise nucleotide diversity for maize (blue) and teosinte (orange).
+Mean values for each taxa are represented by corresponding vertical lines.
 Pairwise nucleotide diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
@@
 We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
-Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively).
-Figure \ref{fig:figure3}C shows the distribution of pairwise nucleotide diversity in each window within our region of interest and each vertical line represents the number of windows with the corresponding value of $\pi$.
+Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.

### 2016-01-30 cover_letter.tex
updating cover letter

--- 2015-12-28 cover_letter.tex
@@
-ANGSD-wrapper is intended to simplify analysis and visualization of next generation resequencing data.
+Dear Managing Editor,
 
-Visualization of analysis results is implemented using Shiny graphics in R.
+Enclosed you will find our manuscript, "ANGSD-wrapper: utilities for analyzing next generation sequencing data." This manuscript reports the novel software, "ANGSD-wrapper," which simplifies analysis and visualization of next generation resequencing data.
+The focus of this program is on usability and it extends utlity of the previously published software, "ANGSD" and programs related to it.
+
+ANGSD-wrapper consists of two parts: 1) ANGSD pipelines that simplify multi-step analyses into a single step, and 2) interactive graphing based on the output of ANGSD-wrapper.
+Interactive visualization of results is implemented using Shiny graphics in R, which uses modern web technologies such as Javascript and HTML5.
 
 The majority of the code is written in Bash, with graphics implemented in R. For speed, one sorting routine is implemented in Python.
+We demonstrate the usefulness of the software by analyzing previously published data in maize and teosinte.
+
+The usability and usefulness of our software should appeal to the readership of Molecular Ecology Resources.
+
+Suggested reviewers for the manuscript include:
+	- Sung Chun; Brigham and Women’s Hospital and Harvard Medical School; sgchun@partners.org 
+	- Naoki Takebayashi; University of Alaska, Fairbanks; ntakebayashi@alaska.edu
+
+Thank you,
+
+Arun Durvasula and Jeffrey Ross-Ibarra

### 2016-01-30 manuscript.tex
edited manuscript.tex

@@
 % \cl{added text}
-ANGSD uses probabilistic approaches to calculate genome wide descriptive statistics.
-The package is specifically designed to produce more accurate results for samples with low sequencing depth and makes use of full genome data while handling a wide array of sampling and experimental designs.
+ANGSD uses probabilistic approaches to calculate genome-wide descriptive statistics.
+The package makes use of genotype likelihood estimates rather than SNP calls and is specifically designed to produce more accurate results for samples with low sequencing depth.
+ANGSD makes use of full genome data while handling a wide array of sampling and experimental designs.
 Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
 % \cl{added text}
-This software supports multiple types of analyses such as Site Frequency Spectrum, Genotype Likelihood estimations, Thetas diversity statistics and neutrality tests, Principal Component Analysis, Admixture, and Inbreeding coefficient calculations.
+ANGSD-wrapper software supports multiple types of analyses including providing estimates of nucleotide sequence diversity and performing neutrality tests.
+ANGSD-wrapper also provides support for Principal Component Analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression.
 %ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
@@
 
-High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or survey fine-scale patterns of variation across the genome. 
+High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or fine-scale variation across the genome.
 Application of these approaches has led to changes in both experimental design and data analysis \citep{ekblom2011applications}.
@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample. 
-This is especially true of studies that span multiple subpopulations, as this balance must be met for each partition \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While the experimental design challenges with molecular population genetic studies have existed for at least a decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
+This is especially true of studies that span multiple populations, as this balance must be met for each partition of the sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
+While the experimental design challenges in molecular population genetic studies have existed for at least two decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
 
@@
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
+% \plm{this would really benefit from a citation}
+% \plm{below I'm confused as to how quotes names should appear. We have both `word' and ``word''.}
 
-Here we present ANGSD-wrapper, a user-friendly interface to stable versions of ANGSD. 
-ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
+Here we present ANGSD-wrapper, a user-friendly interface to ANGSD. 
+ANGSD-wrapper takes the form of a set of configuration files and `wrapper` scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
 ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}.
@@
 % \ph{This could be reworded} 
-Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
-Additionally, the wrapper scripts are written against `frozen' versions of ANGSD and supporting tools, for consistency of analysis. 
+Furthermore, ANGSD-wrapper permits whole genome analysis of population genetic data or the potential to focus a set genomic regions.
+The wrapper scripts are written against `frozen` versions of ANGSD and supporting tools, for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
-We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
+We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage, whole-genome data from domesticated maize and two related wild teosinte subspecies. 
 ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
@@
 An installation of the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.  
-The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor} for the visualization package.
+The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor}.
 These dependent packages are installed automatically upon first run of visualization interface.
@@
 % \ph{Necessary?} \ad{changed the wording}
-The visualization software included with ANGSD-wrapper is contained in a directory called ``shinyGraphing.''
+The visualization software included with ANGSD-wrapper is contained in a directory called `shinyGraphing.`
 This application is started from within ANGSD-wrapper and launched locally from a standard web browser. 
@@
 distribution of pairwise nucleotide diversity for maize (blue) and teosinte (orange).
-Mean values for each taxa are represented by corresponding vertical lines.
+Mean values for each taxon are represented by corresponding vertical lines.
 Pairwise nucleotide diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
@@
 {\it parviglumis}, and {\it Zea mays} ssp.
-{\it mexicana} (\textit{mex}) with colors representing the K=4 source populations.
+{\it mexicana} (\textit{mex}) with colors representing the {\it K}=4 source populations.
 Individuals are shown in the same order as Table \ref{tab:samples}. }
@@
 This analysis is a supervised method that assigns proportions of the each individual's genome to a putative source population, represented by the colors. 
-We ran this analysis over a range of number of ancestral populations (K= 2, 3, 4; See Figure \ref{fig:suppadmix}) and chose the value with the maximum likelihood while keeping the number of parameters as low as possible.
+We ran this analysis over a range of number of ancestral populations ({\it K}= 2, 3, 4; See Figure \ref{fig:suppadmix}) and chose the value with the maximum likelihood while keeping the number of parameters as low as possible.
 In maize, we see that most individuals come from one source population (dark grey or dark blue), which is evidence for no admixture between these lowland maize samples and  ssp.
@@
 A single ssp.
-\textit{parviglumis} accession from the Northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
+\textit{parviglumis} accession from the northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
+% \plm{"does not appear to be well-classified with these data is very vague"}
 
@@
 ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.  
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM), from a University of Minnesota Doctoral Dissertation Fellowship supporting TJYK, and from the University of California Davis Plant Sciences Department.

### 2016-01-31 manuscript.tex
jri edits

@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
-% \cl{added text}
 ANGSD uses probabilistic approaches to calculate genome-wide descriptive statistics. 
@@
 ANGSD makes use of full genome data while handling a wide array of sampling and experimental designs. 
-Here we present ANGSD-wrapper, a user-friendly interface for running ANGSD and visualizing results.
-% \cl{added text}
-ANGSD-wrapper software supports multiple types of analyses including providing estimates of nucleotide sequence diversity and performing neutrality tests.
-ANGSD-wrapper also provides support for Principal Component Analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression.
-%ANGSD-wrapper includes a number of 'wrapper' scripts that facilitate configuration and execution of multi-step analyses.
+Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
+ANGSD-wrapper supports multiple types of analyses including estimates of nucleotide sequence diversity and performing neutrality tests, principal component analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression.
 ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
@@
 
-High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals or fine-scale variation across the genome.
+High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals and interrogate genome-wide variation. 
 Application of these approaches has led to changes in both experimental design and data analysis \citep{ekblom2011applications}. 
 Many popular software packages used by researchers for analysis of comparative resequencing data \citep[see][]{excoffier2006computer} were not designed to handle these novel data types or efficiently analyze the large volumes of data now being generated.
-Despite the decreasing cost of sequencing, researchers must allocate finite resources and balance the depth of sequencing and the breadth of a sample. 
-This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample. 
-This is especially true of studies that span multiple populations, as this balance must be met for each partition of the sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While the experimental design challenges in molecular population genetic studies have existed for at least two decade, high throughput sequencing brings the technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
+Despite the decreasing cost of sequencing, researchers must nonetheless allocate finite resources and balance the depth of sequencing and the breadth of a sample. 
+This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
+While these experimental design challenges in molecular population genetic studies have existed for at least two decade, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
 
 A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}.
- \citet{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
+ \citet{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
 ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
@@
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage.
-% \plm{this would really benefit from a citation}
-% \plm{below I'm confused as to how quotes names should appear. We have both `word' and ``word''.}
 
 Here we present ANGSD-wrapper, a user-friendly interface to ANGSD. 
-ANGSD-wrapper takes the form of a set of configuration files and `wrapper` scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
+ANGSD-wrapper takes the form of a set of configuration files and wrapper scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
 ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
-%\tjyk{Fumagali citation doesn't show up properly}\ph{I think the citation is just 'Fumagali' and the link to the webpage.
-If so, then it works for me -PH} \ad{yes, there's no citation other than the github repo (and no DOI there)}
-% \ph{This could be reworded} 
-Furthermore, ANGSD-wrapper permits whole genome analysis of population genetic data or the potential to focus a set genomic regions.
-The wrapper scripts are written against `frozen` versions of ANGSD and supporting tools, for consistency of analysis. 
+As in ANGSD, ANGSD-wrapper allows users to perform whole genome analysis or analyze a set of user-defined windows across the genome.
+The wrapper scripts are written against a frozen versions of ANGSD and supporting tools for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
-We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage, whole-genome data from domesticated maize and two related wild teosinte subspecies. 
+We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage resequencing using data from domesticated maize and two related wild teosinte subspecies. 
 ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
@@
 The scripts can be run either on a standalone computer with a UNIX terminal, or on computing clusters where they can be submitted to a queuing system such as SGE \citep{Microsystems):2001:SGE:560889.792378}, Slurm \citep{Jette02slurm:simple} or TORQUE \citep{Staples:2006:TRM:1188455.1188464}.
-% \ph{has anyone tested AW on a cluster?} \ad{yep I have.
-I logged into a compute node via srun and ran it that way}
-A Python installation \citep{Python} is required for some light, dynamic pre-processing of the data.
-An installation of the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.  
+A Python installation \citep{Python} is required for some light, dynamic pre-processing of the data, and the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.
 The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor}.
-These dependent packages are installed automatically upon first run of visualization interface.
+These dependent packages are installed automatically upon first run of the visualization interface.
 
@@
 
-% \ph{Necessary?} \ad{changed the wording}
 The visualization software included with ANGSD-wrapper is contained in a directory called `shinyGraphing.`
@@
 Each tab in the GUI contains plots for different ANGSD methods.
-
 In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate results file. 
-The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.
+The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built-in image saving capabilities.
 
@@
 \end{figure}
-
-\section*{Results}
 
@@
 
-%\tjyk{Should this start a "Results" section?} \ad{yes I think so.}
-
+\section*{Results}
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of pairwise nucleotide diversity in a single genomic region of domesticated maize and wild teosinte. 
 We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
-The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442}. 
+The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442} and can be downloaded from within AGNSD-wrapper using teh command
+
+\begin{lstlisting}
+$ angsd-wrapper setup data
+\end{lstlisting}
+
 In the following we refer to methods listed in Table \ref{tab:methods} when describing analyses.
@@
 Using the ``2DSFS method,'' which includes an \fst calculation, we find a mean \fst in this region of 0.116, nearly identical to the  genome-wide value of 0.11 reported in \cite{hufford2012comparative}.
-There are no genes in this region that have  been identified as potential domestication candidates, consistent with the lack of extended regions of high \fst in our analysis (Figure \ref{fig:suppfst}).
+None of the genes in this region have been identified as potential domestication candidates \cite{hufford2012comparative}, consistent with the lack of extended regions of high \fst in our analysis (Figure \ref{fig:suppfst}).
 
@@
 {\it parviglumis}, and {\it Zea mays} ssp.
-{\it mexicana} (\textit{mex}) with colors representing the {\it K}=4 source populations.
-Individuals are shown in the same order as Table \ref{tab:samples}. }
+{\it mexicana} (\textit{mex}).
+Colors represent the proportion of each individual's genome assigned to one of the {\it K}=4 source populations.
+Individuals are shown in the same order as Table \ref{tab:samples}. For results using other values of \texit{K} see Figure \ref{fig:suppadmix}.}
 \label{fig:admixture}
 \end{figure}
-% \jri{what are vertical lines in pi density plot?
-need to be consistent about saying "nucleotide diversity" vs "pairwise diversity" vs "pairwise differences" etc.
-I vote for "pairwise nucleotide diversity"} \ad{changed.}
+
 Finally, we include two samples of the related wild teosinte \textit{Zea mays} ssp. \textit{mexicana} to assess evidence for admixture.  
-Using the ``Admixture'' method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}, we identify structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}).
-This analysis is a supervised method that assigns proportions of the each individual's genome to a putative source population, represented by the colors. 
-We ran this analysis over a range of number of ancestral populations ({\it K}= 2, 3, 4; See Figure \ref{fig:suppadmix}) and chose the value with the maximum likelihood while keeping the number of parameters as low as possible.
-In maize, we see that most individuals come from one source population (dark grey or dark blue), which is evidence for no admixture between these lowland maize samples and  ssp.
-\textit{mexicana}, and is consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}.
+We ran the ``Admixture'' method, which implements an estimate of admixture proportion from genotype likelihoods \citep{pmid24026093}.
+We identify structure within domesticated maize separating three high-latitude temperate landraces from the other tropical accessions (Figure \ref{fig:admixture}). 
 \textit{Zea mays} ssp.
@@
 \textit{parviglumis} collected from a region in which many teosinte populations appear to be the result of admixture between the two subspecies \citep{fang2012megabase}.
-A single ssp.
-\textit{parviglumis} accession from the northern extent of the range does not appear to be well-classified with these data, likely due to our relatively limited geographic and genomic sampling.
-% \plm{"does not appear to be well-classified with these data is very vague"}
+Consistent with an independent analysis from SNP genotyping \citep{hufford2013genomic}, the lowland maize samples included here show no evidence of admixture with ssp. \textit{mexicana}.
+Most ssp.
+\textit{parviglumis} accessions fall primarily into a single (light blue) cluster, but two accessions show assignment to multiple clusters, perhaps due to the limited resolution resulting from analysis of a single genomic region and restricted geographic sampling.
 

### 2016-02-01 manuscript.tex
final final final

@@
 \usepackage{pgf}
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=2.5cm, top=3cm]{geometry}
 \usepackage[normalem]{ulem}
@@
             citecolor=gray]{hyperref}
+\usepackage{setspace}
+\usepackage{lineno}
+\linenumbers
+\doublespacing
 
@@
 
+Running head: ANGSD-wrapper
+
+Keywords: software, visualization, \textit{Zea}, domestication, population genetics
+
+\pagebreak
+
 \begin{abstract}
@@
 \begin{center}
+    \caption{Table of methods implemented in ANGSD-wrapper}
     \begin{tabular}{ | p{3.5cm} | p{5cm} | p{2.5cm}  |}
@@
     \end{tabular}
-    \caption{Table of methods implemented in ANGSD-wrapper}
     \label{tab:methods}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
 \caption{Visualization of Watterson's $\theta$ estimated by ANGSD across a 1.5Mb region of chromosome 10 in {\it Zea mays} ssp.
@@
 \label{fig:theta} }
+\includegraphics[width=0.8\linewidth]{figures/fig1.pdf}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/figure3big.pdf}
 \caption{Summary statistics for {\it Zea mays}.
@@
 Pairwise nucleotide diversity results are visualized separately from the interactive graphics and colors were added to A and B using a custom script.}
+\includegraphics[width=0.8\linewidth]{figures/figure3big.pdf}
 \label{fig:figure3}
@@
 
-\section*{Results}
+\section*{Results and Discussion}
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of pairwise nucleotide diversity in a single genomic region of domesticated maize and wild teosinte. 
@@
 \centering
-\includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
 \caption{Admixture analysis for {\it Zea mays} ssp.
@@
 Individuals are shown in the same order as Table \ref{tab:samples}. For results using other values of \texit{K} see Figure \ref{fig:suppadmix}.}
+\includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}
 \label{fig:admixture}
@@
 \clearpage
+\singlespacing
 \bibliographystyle{apa}

### 2016-04-28 manuscript.tex
final^5
fixed typos

@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While these experimental design challenges in molecular population genetic studies have existed for at least two decade, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
+While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
 
@@
 ANGSD-wrapper takes the form of a set of configuration files and wrapper scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
+ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsTools \citep{fumagalli2014ngstools}, ngsF\citep{vieira2013estimating}, ngsAdmix \citep{pmid24026093}, and PCA \citep{fumagalli2013quantifying}.
 As in ANGSD, ANGSD-wrapper allows users to perform whole genome analysis or analyze a set of user-defined windows across the genome.
-The wrapper scripts are written against a frozen versions of ANGSD and supporting tools for consistency of analysis. 
+The wrapper scripts are written against a frozen versions of ANGSD (v [INSERT VERSION]) and supporting tools for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
@@
 ANGSD-wrapper is divided into scripts associated with analytical approaches implemented in ANGSD and associated software. 
-ANGSD-wrapper provides a common configuration file, \texttt{Common\_Config}, which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
+It provides a common configuration file, \texttt{Common\_Config}, which holds variables that are likely to remain constant across analyses, including identifiers for chromosomal regions and the paths to project directories.
 In ANGSD-wrapper, each method is self-contained in a shell script which uses information from the common configuration file and a method-specific configuration file. 
@@
 We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
-The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442} and can be downloaded from within AGNSD-wrapper using teh command
+The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442} and can be downloaded from within ANGSD-wrapper using the command
 
@@
 \textit{parviglumis}, assuming an inbreeding coefficient of $F=1$ for these highly inbred samples.
-The SFS and diversity statistics were calculated using ancestral states inferred in the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
+The SFS and diversity statistics were calculated using ancestral states inferred by the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
 We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
@@
 Colors represent the proportion of each individual's genome assigned to one of the {\it K}=4 source populations.
-Individuals are shown in the same order as Table \ref{tab:samples}. For results using other values of \texit{K} see Figure \ref{fig:suppadmix}.}
+Individuals are shown in the same order as Table \ref{tab:samples}. For results using other values of \textit{K} see Figure \ref{fig:suppadmix}.}
 \includegraphics[width=0.8\linewidth]{figures/mt4labeled.pdf}

### 2016-04-29 manuscript.tex
added examples of ANGSD use in intro

@@
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
+Further, this probablistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or funding prevents the ability to obtain high quality data. 
+Additionally, for many of these non-model organisms, existing tools are not able to take into account biological differences such as non-random mating into their statistical models.
+Recent use of ANGSD in the study of non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} has allowed population genetic analysis despite using low coverage resequencing data.
 

### 2016-04-29 manuscript.tex
typo

@@
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
-Further, this probablistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or funding prevents the ability to obtain high quality data. 
+Further, this probabilistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or funding prevents the ability to obtain high quality data.
 Additionally, for many of these non-model organisms, existing tools are not able to take into account biological differences such as non-random mating into their statistical models.

### 2016-05-02 manuscript.tex
jri minor edits

@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
-ANGSD uses probabilistic approaches to calculate genome-wide descriptive statistics. 
-The package makes use of genotype likelihood estimates rather than SNP calls and is specifically designed to produce more accurate results for samples with low sequencing depth. 
-ANGSD makes use of full genome data while handling a wide array of sampling and experimental designs. 
+Rather than require users to first call SNPs, however, ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, taking advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth
 Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
@@
  \citet{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
-ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
-One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
+ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format.
+One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on polymorphic sites identified \emph{a priori}. 
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
@@
 Additionally, for many of these non-model organisms, existing tools are not able to take into account biological differences such as non-random mating into their statistical models.
-Recent use of ANGSD in the study of non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} has allowed population genetic analysis despite using low coverage resequencing data.
+Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{beissinger2015recent}.
 
@@
 A Python installation \citep{Python} is required for some light, dynamic pre-processing of the data, and the statistical software R \citep{Rcitation} is required to make use of the visualization tools incorporated in ANGSD-wrapper.
-The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor}.
-These dependent packages are installed automatically upon first run of the visualization interface.
+The visualization portion of ANGSD-wrapper also requires installation of the R packages \texttt{Shiny} \citep{shiny}, \texttt{APE} \citep{APE}, \texttt{Lattice} \citep{lattice}, \texttt{Hmisc} \citep{Hmisc}, \texttt{data.table} \citep{datatable}, \texttt{DT} \citep{javascriptDT}, and \texttt{shinythemes} \citep{shinythemes}, as well as \texttt{genomeIntervals} \citep{genomeIntervals} from Bioconductor \citep{bioconductor}; these are installed automatically upon first run of the visualization interface.
 
@@
 Analyses supported by ANGSD-wrapper are shown in Table \ref{tab:methods}.
-A detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional details, documentation, a tutorial, and a wiki can be found on the GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
+A detailed flowchart of each of these workflows is shown in Figure \ref{fig:supp2}, and additional  documentation, a tutorial, and a wiki can be found on the project GitHub page: \url{https://github.com/mojaveazure/angsd-wrapper/wiki}.
 
@@
 In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate results file. 
-The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built-in image saving capabilities.
+The Shiny server automatically parses standard ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built-in image saving capabilities.
 
@@
 As a demonstration of analyses in ANGSD-wrapper, we explore patterns of pairwise nucleotide diversity in a single genomic region of domesticated maize and wild teosinte. 
-We used a subset of the resequenced samples from the Maize HapMap2 project (Table \ref{tab:samples}) and calculated summary statistics using a 10Mb region on chromosome 10 \citep{chia2012maize}. 
+We used a subset of the resequenced samples from the Maize HapMap2 project \citep{chia2012maize} and calculated summary statistics using a 10Mb region on chromosome 10 (Table \ref{tab:samples}). 
 The data are available at \url{https://figshare.com/articles/Example_Data_tar_bz2/2063442} and can be downloaded from within ANGSD-wrapper using the command
@@
 The SFS and diversity statistics were calculated using ancestral states inferred by the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
+We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, beissinger2015recent}.
 Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
@@
 ANGSD-wrapper provides an easy-to-use interface that simplifies many population genetic analyses implemented in ANGSD \citep{korneliussen2014angsd} and permits the exploration of genome-scale results through interactive visualization.
-ANGSD-wrapper is under active development to incorporate updates to the ANGSD software package.  
+ANGSD-wrapper is under active development to incorporate new analyses and updates to the ANGSD software package.
+
 \section*{Acknowledgments}
-We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM), from a University of Minnesota Doctoral Dissertation Fellowship supporting TJYK, and from the University of California Davis Plant Sciences Department.
+We acknowledge funding support from the US National Science Foundation (IOS-1238014 to JRI and IOS-1339393 to PLM), from a University of Minnesota Doctoral Dissertation Fellowship supporting TJYK, and from USDA Hatch project CA-D-PLS-2066-H, and from the University of California Davis Plant Sciences Department.
 We thank members of the Ross-Ibarra and Morrell Labs for discussion and software testing.

### 2016-05-15 manuscript.tex
PLM edits to text

@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
-Rather than require users to first call SNPs, however, ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, taking advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth
+ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, thus SNP calling is not a required for comparative analyses.
+This takes advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth.
 Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
-ANGSD-wrapper supports multiple types of analyses including estimates of nucleotide sequence diversity and performing neutrality tests, principal component analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression.
+ANGSD-wrapper supports multiple types of analyses including estimates of nucleotide sequence diversity and performs neutrality tests, principal component analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression.
 ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
@@
 
-High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals and interrogate genome-wide variation. 
+High throughput sequencing has revolutionized evolutionary genetics, allowing researchers to quickly assay large numbers of individuals and examine genome-wide variation. 
 Application of these approaches has led to changes in both experimental design and data analysis \citep{ekblom2011applications}. 
@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates.
+While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates \citep{}.
 
 A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}.
- \citet{korneliussen2014angsd} recently published the software package ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
+The software package ANGSD \citet{korneliussen2014angsd}, is noteworthy because it enables users to perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}.
 ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format.
-One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on polymorphic sites identified \emph{a priori}. 
+One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than on predetermined polymorphic sites \emph{a priori}. 
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
-Further, this probabilistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or funding prevents the ability to obtain high quality data.
-Additionally, for many of these non-model organisms, existing tools are not able to take into account biological differences such as non-random mating into their statistical models.
+Further, this probabilistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or cost considerations prevent the ability to obtain high quality data.
+Also, many existing tools do not implement statistical models that account for biological issues such as non-random mating.
 Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{beissinger2015recent}.
@@
 ANGSD-wrapper takes the form of a set of configuration files and wrapper scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
-ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsTools \citep{fumagalli2014ngstools}, ngsF\citep{vieira2013estimating}, ngsAdmix \citep{pmid24026093}, and PCA \citep{fumagalli2013quantifying}.
+ANGSD-wrapper also assists with configuration of ANGSD-related programs such as ngsTools \citep{fumagalli2014ngstools}, ngsF\citep{vieira2013estimating}, ngsAdmix \citep{pmid24026093}, and PCA \citep{fumagalli2013quantifying}.
 As in ANGSD, ANGSD-wrapper allows users to perform whole genome analysis or analyze a set of user-defined windows across the genome.
@@
 In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate results file. 
-The Shiny server automatically parses standard ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built-in image saving capabilities.
+The Shiny server automatically parses standard ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built-in image export options.
 
@@
 The SFS and diversity statistics were calculated using ancestral states inferred by the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-We show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, beissinger2015recent}.
+We show that the maize SFS is skewed toward intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, beissinger2015recent}.
 Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).
@@
 We thank the authors of ANGSD and related programs for answering questions, particularly Matteo Fumagalli and Filipe Vieira.
+Matteo Fumagalli and an anonymous reviewer provided comments that improved both the manuscript and the implementation of ANGSD-wrapper.
 Finally, we would like to thank Felix Andrews for statistical advice, although we did not follow it.

### 2016-05-16 manuscript.tex
changed supp figure and caption

@@
 Also, many existing tools do not implement statistical models that account for biological issues such as non-random mating.
-Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{beissinger2015recent}.
+Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{Beissinger031666}.
 
@@
 As in ANGSD, ANGSD-wrapper allows users to perform whole genome analysis or analyze a set of user-defined windows across the genome.
-The wrapper scripts are written against a frozen versions of ANGSD (v [INSERT VERSION]) and supporting tools for consistency of analysis. 
+The wrapper scripts are written against a frozen versions of ANGSD (v0.902-48-g8b89ba4) and supporting tools for consistency of analysis. 
 Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.
@@
 The SFS and diversity statistics were calculated using ancestral states inferred by the ``Ancestral Sequence'' method from a single resequenced genome of \textit{Tripsacum dactyloides}.  
-We show that the maize SFS is skewed toward intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, beissinger2015recent}.
+We show that the maize SFS is skewed toward intermediate-frequency variants (Figure \ref{fig:figure3}A-B, Tajima's D of 0.2 and 0.0085 in maize and teosinte, respectively), likely a result of the bottleneck associated with maize domestication \citep{eyre1998investigation, Beissinger031666}.
 Using the ``Thetas'' method we find further evidence of the domestication bottleneck, with mean levels of pairwise nucleotide diversity in this region in maize $\approx 25\%$ lower than in teosinte (0.0061 and 0.0082, respectively; Figure \ref{fig:figure3}C).

### 2016-05-16 manuscript.tex
small changes

@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
-ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, thus SNP calling is not a required for comparative analyses.
+ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, thus SNP calling is not required for comparative analyses.
 This takes advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth.
@@
 ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format.
-One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than on predetermined polymorphic sites \emph{a priori}. 
+One of the most important features of ANGSD is that analyses can be integrated over per-site genotype likelihoods, rather than on predetermined polymorphic sites \emph{a priori}. 
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 

### 2016-05-23 manuscript.tex
minor last edits

--- 2016-05-16 manuscript.tex
@@
 The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
-ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, thus SNP calling is not a required for comparative analyses.
-This takes advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth.
+ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, and SNP calling is thus not required for comparative analyses.
+This takes advantage of all of the sequencing data and produces more accurate results for samples with low sequencing depth.
 Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
@@
 ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format.
-One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than on predetermined polymorphic sites \emph{a priori}. 
+One of the most important features of ANGSD is that analyses can be integrated over all sites, rather than on predetermined polymorphic sites \emph{a priori}. 
 This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 

### 2016-05-29 manuscript.tex
Attempted to add citations...

@@
 \affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN 55108}
-\affil[3]{Center for Population Biology and Genome Center, University of California, Davis, CA 95616}
+\affil[3]{Center for Population Biology and Genome Center, University of lynch2008diversityCalifornia, Davis, CA 95616}
 \affil[$\dag$]{These authors contributed equally.}
@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates \citep{}.
+While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates \citep{lynch2008diversity, lynch2008diversity}.
 

### 2016-05-30 manuscript.tex
fixed manuscript

--- 2016-05-23 manuscript.tex
@@
 This poses a challenge for population genetics analysis, which generally requires accurate polymorphism calls in a broad sample \citep{pluzhnikov1996optimal, felsenstein2006accuracy}. 
-While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates \citep{}.
+While these experimental design challenges in molecular population genetic studies have existed for at least two decades, high throughput sequencing brings the added technical challenges of highly variable coverage, missing data, and high per-nucleotide error rates \citep{lynch2008diversity, johnson2007bias}.
 
